\section{Ordinary differential equations}

In this chapter the integrand is taken as a path in \(\R^e\), note that it could have been a path in a finite dimensional Banach space.
Later the space \( T^N(\R^d)\) will be an example.

We consider equations of the type:
\begin{equation}
    dy = V(y) dx,
\end{equation}
where $y$ is a path in $\R^e,$ $x$ is a path in $\R^d$ and $V:\R^e \rightarrow L(\R^e, \R^d).$
Observe that $V$ can be seen as $V = (V_1, \ldots, V_d),$ where $V_i: \R^e \rightarrow \R^e.$
One can see the $V_i(y)$s as the columns of $V(y)$ seen as a matrix.
In this case:
\begin{equation}
    \int V(y) dx = \sum_i^d \int V_i(y) dx_i.
\end{equation} 

\begin{definition}
    A collection of vector fields $V = (V_1, \ldots, V_n) : \R^e \rightarrow L(\R^e, \R^d)$ is said \textit{bounded} if
    \begin{equation}
        |V|_\infty = sup_{y \in \R^e} |V(y)| < \infty.
    \end{equation}

    For any $U \subset \R^e,$ we define the $1$-Lipschitz norm as:
    \begin{equation}
        |V|_{Lip^1(U)} := \max \left\{ sup_{y \neq z,\ y,z \in U} \frac{|V(y) - V(z)|}{|y - z|}, sup_{y \in U} |V(y)| \right\}.
    \end{equation}

    We say that $V \in Lip^1(\R^e)$ if $|V|_{Lip^1(\R^n)} < infty$ and locally 1-Lipschitz if it is 1-Lipschitz in any bounded subset $U$ of $\R^e.$
\end{definition}

\begin{lemma}[Gronwall's lemma]
    Let $x \in C^{1-var}([0,T], \R^d)$ and $\phi:[0,T] \rightarrow \R^+$ a bounded measurable function. Assume that for all $t \in [0,T]$
    \begin{equation}\label{eq:gronwall}
        \phi_t \leq K + L \int_0^t \phi_s |dx_s|.
    \end{equation}

    Then for all $t \in [0,T]$
    \begin{equation}
        \phi_t \leq K exp\left( L \int_0^t |dx_s| \right).
    \end{equation}

    The same works substituting $K$ with a non-decreasing function $K_t.$
\end{lemma}
\begin{proof}[Outline of the proof]
    Apply inductively Equation \eqref{eq:gronwall} and use the exponential series definition.
    It also requires the iterated integral of $|dx_t|,$ which is easily solved with integration by parts.
\end{proof}

\subsection{Existence}

\begin{theorem}[Existence]
    Assume that
    \begin{itemize}
        \item $V = (V_1, \ldots, V_d)$ is a collection of continuous, bounded vector fields on $\R^e$,
        \item $y_0 \in \R^e$ is the initial condition,
        \item $x \in C^{1-var}([0,T], \R^d)$.
    \end{itemize}
    Then the exist a solution (not necessarily unique) of the ODE $dy = V(y)dx$.
    Moreover for all $s < t$,
    \begin{equation}
        |\pi_{(V)}(0, y_0; x)|_{1-var, [s,t]} \leq |V|_\infty |x|_{1-var, [s,t]}.
    \end{equation}
\end{theorem}
\begin{proof}[Outline of the proof]
    Define the Euler approximation on a dissection $D$.
    Bound the variations $y^D_{s,t}$ and consequently the $|y^D|_\infty.$
    Then by Ascoli Arzelà $y^{D_n}$ must have a limit point.
    Hence one can pass to the limit in the integral that defined $y^D_{s,t}$ by dominated convergence.
    Note that one has to pass to the limit in $V(y_{r_D}^D)$, where $r_D$ is the largest element in $D$ smaller than $r.$
\end{proof}

\begin{theorem}[Existence up to explosion time]
      Assume that
    \begin{itemize}
        \item $V = (V_1, \ldots, V_d)$ is a collection of continuous vector fields on $\R^e$,
        \item $y_0 \in \R^e$ is the initial condition,
        \item $x \in C^{1-var}([0,T], \R^d)$.
    \end{itemize}
    Then either there exists a solution (not necessarily unique) of the ODE $dy = V(y)dx$, $y:[0,T] \rightarrow \R^e,$
    or there exist a $\tau \in [0,T]$ and a local solution $y:[0,\tau) \rightarrow \R^e$ that solves in any $[0,t], \ t < \tau,$ and 
 \begin{equation}
    lim_{t \nearrow \tau} |y_t| = +\infty.
 \end{equation}
\end{theorem}
\begin{proof}[Outline of the proof]
    Consider the solution in balls around $y_0$ of increasing radius. Up to reaching the border at $\tau_i$ of each ball we have a solution by the Existence theorem.
    Proceeding inductively we either reach $T$ with the $\tau_i$ or we converge to a $\tau.$ In the second case there is explosion.
\end{proof}

\begin{theorem}[No explosion]
 Assume that
    \begin{itemize}
        \item $V = (V_1, \ldots, V_d)$ is a collection of continuous vector fields on $\R^e$ of \textbf{linear growth}, that is
        \begin{equation}\label{eq:lineargrowth}
             \exists A > 0:\ |V(y)| < A (1 + |y|),\ \forall y \in \R^e,
        \end{equation}
        \item $y_0 \in \R^e$ is the initial condition,
        \item $x \in C^{1-var}([0,T], \R^d)$ and $l \geq A |x|_{1-var,[0,T]}.$
    \end{itemize}   
    Then explosion cannot happen. Moreover any solution satisfies:
    \begin{equation}
        |y|_{\infty, [0,T]} \leq (|y_0| + l) \exp(l).
    \end{equation}

    Furthermore for all $0 \leq s \leq t \leq T,$ 
    \begin{equation}
        |y|_{1-var, [s,t]} \leq (1 + |y_0|) \exp(2l) A \int_s^t |dx_u|.
    \end{equation}
\end{theorem}
\begin{proof}[Outline of the proof]
    Use the condition in Equation \eqref{eq:lineargrowth} plus Gronwall to control $|y_s|.$
    Then use Gronwall again to control $|y_{u,v}|.$
\end{proof}

\subsection{Uniqueness}

\begin{theorem}
    Assume that 
    \begin{itemize}
        \item $V = (V_1, \ldots, V_d)$ is a collection of Lipschitz continuous vector fields on $\R^e,$ such that
        \begin{equation}
            v \geq \sup_{y,z \in \R^e} \frac{|V(y) - V(z)|}{|y - z|}.
        \end{equation}
        \item $x \in C^{1-var}([0,T],\R^d)$ such that $v |x|_{1-var,[0,T]} \geq l.$
    \end{itemize}
    Then, for every initial condition $y_0$ there exists a unique ODE solution and the associated flow is Lipschitz continuous, that is:
    \begin{equation}\label{eq:solutionLipschitz}
         |\pi_{(V)}(0, y^1_0; x) - \pi_{(V)}(0, y^2_0; x)|_\infty \leq |y^1_0 - y^2_0| \exp(l).
    \end{equation}
    Moreover, for all $s < t$ in $[0,T],$
    \begin{equation}
        |\pi_{(V)}(0, y^1_0; x) - \pi_{(V)}(0, y^2_0; x)|_{1-var,[s,t]} \leq |y^1_0 - y^2_0| \exp(2l) v |x|_{1-var,[s,t]}.
    \end{equation}
\end{theorem}
\begin{proof}[Outline of the proof]
    First use Gronwall on $\bar{y} = y^1 - y^2.$ This is enough to prove Equation \eqref{eq:solutionLipschitz}.
    Then use Gronwall again on $\bar{y}_{s,t}.$
\end{proof}

\subsection{Consequences of uniqueness}
\begin{proposition}[Change of variable]
    Let $x \in C^{1-var}([0,T_1], \R^d)$ and $V = (V_1, \ldots, V_d)$ a collection of locally Lipschitz vector fields on $\R^e$ of linear growth.
    Assume $\phi$ is a non-decreasing function from $[0,T_2]$ onto $[0,T_1].$ so that
\begin{equation}
    x \circ \phi \in C^{1-var}([0,T_2], \R^d).
\end{equation}
Then 
\begin{equation}
    \pi_{(V)}(0, y_0; x)_{\phi(\cdot)} = \pi_{(V)}(0, y_0; x \circ \phi) \text{ on } [0,T_2].
\end{equation}
\end{proposition}

\subsection{Continuity of the solution map}

\subsubsection{Limit theorem for 1-variation signals}

\begin{theorem}
    We consider
    \begin{itemize}
        \item $V^1 = (V_1^1, \ldots V^1_d)$ and $V^2 = (V_2^1, \ldots V^2_d)$ two collections of $Lip^1$ vector fields on $\R^e$ such that
        \begin{equation}
            \max_{i=1,2} |V^i|_{Lip^1} \leq v.
        \end{equation}
        \item $y_0^1$ and $y_0^2$ initial conditions.
        \item $x^0$ and $x^1$ are paths in $C^{1-var}([0,T], \R^d)$ such that 
        \begin{equation}
            \max_{i=1,2} |x^i|_{1-var, [0,T]} \leq l.
        \end{equation} 
    \end{itemize}
    Then, if $y^i = \pi_{V_i}(0, y_0^i; x^i)$ we have\footnote{Recall that $|x|_0 = \sup_{u,v \in [0,T]} |x_u - x_v|$}.
    \begin{equation}
        |y^1 - y^2|_\infty \leq \left( |y_0^1 - y_0^2| + v |x^1 - x^2|_0 + |V^1 - V^2|_\infty l \right) \exp(2vl).
    \end{equation}
\end{theorem}
\begin{proof}[Outline of the proof]
    Add and subtract, integration by parts and finally Gronwall.
\end{proof}

\begin{corollary}[Localization]
      We consider
    \begin{itemize}
        \item $V^1 = (V_1^1, \ldots V^1_d)$ and $V^2 = (V_2^1, \ldots V^2_d)$ two collections of locally $Lip^1$ vector fields on $\R^e$ with linear growth,
        \item $y_0^1$ and $y_0^2$ initial conditions, with $R \geq |y_0^i|,$
        \item $x^0$ and $x^1$ are paths in $C^{1-var}([0,T], \R^d)$ such that 
        \begin{equation}
            \max_{i=1,2} |x^i|_{1-var, [0,T]} \leq l.
        \end{equation} 
    \end{itemize}
    Then, if $y^i = \pi_{V_i}(0, y_0^i; x^i)$
     \begin{equation}
        |y^1 - y^2|_\infty \leq C \left( |y_0^1 - y_0^2| + |x^1 - x^2|_0 + |V^1 - V^2|_{\infty, B(0,M)} \right).
    \end{equation}
\end{corollary}

\begin{exercise}[Change of variable formula]
    Change of variable.
\end{exercise}

\subsubsection{Continuity under the 1-variation norm}

\begin{theorem}
    We consider
    \begin{itemize}
        \item $V^1 = (V_1^1, \ldots V^1_d)$ and $V^2 = (V_2^1, \ldots V^2_d)$ two collections of $Lip^1$ vector fields on $\R^e$ such that
        \begin{equation}
            \max_{i=1,2} |V^i|_{Lip^1} \leq v,
        \end{equation}
        \item $y_0^1$ and $y_0^2$ initial conditions, with $R \geq |y_0^i|,$
        \item $x^1$ and $x^2$ are paths in $C^{1-var}([0,T], \R^d)$ such that 
        \begin{equation}
            \max_{i=1,2} |x^i|_{1-var, [0,T]} \leq l.
        \end{equation} 
    \end{itemize}

    Then, if $y^i = \pi_{V_i}(0, y_0^i; x^i)$
    \begin{equation}
        |y^1 - y^2|_{1-var,[0,T]} \leq 2 \left( |y_0^1 - y_0^2| v l + |x^1 - x^2|_{1-var,[0,T]} + |V^1 - V^2|_\infty l \right) e^{3vl}.
    \end{equation}
\end{theorem}

\begin{corollary}[Localization, the book uses $|\cdot|_0$ on $x$, seems wrong]
        We consider
    \begin{itemize}
        \item $V^1 = (V_1^1, \ldots V^1_d)$ and $V^2 = (V_2^1, \ldots V^2_d)$ two collections of locally $Lip^1$ vector fields on $\R^e,$ with linear growth,
        \item $y_0^1$ and $y_0^2$ initial conditions, with $R \geq |y_0^i|,$
        \item $x^1$ and $x^2$ are paths in $C^{1-var}([0,T], \R^d)$ such that 
        \begin{equation}
            \max_{i=1,2} |x^i|_{1-var, [0,T]} \leq l.
        \end{equation} 
    \end{itemize}

    Then, if $y^i = \pi_{V_i}(0, y_0^i; x^i)$ there exist constants $C, M$ depending only on $R, l$ and the vector fields $V$ such that
    \begin{equation}
        |y^1 - y^2|_{1-var,[0,T]} \leq C \left( |y_0^1 - y_0^2| + |x^1 - x^2|_{1-var,[0,T]} + |V^1 - V^2|_{\infty, B(0,M)} \right).
    \end{equation}
\end{corollary}

\subsection{Summary of the regularity estimates for ODEs}
Summarize in a scheme conditions on $V$ and their consequence on $y.$

\begin{itemize}
    \item \textit{existence}: $V$ continuous, bounded $\Rightarrow$ Global existence and $1-var$ estimate.
    \item \textit{explosion}: $V$ continuous $\Rightarrow$ Existence up to explosion time $\tau$ or global.
    \item \textit{non explosion}: $V$ continuous, linear growth $\Rightarrow$ No explosion, uniform and $1-var$ estimates.
    \item \textit{uniqueness}: $V$ globally $Lip^1$ $\Rightarrow$ esistence and uniqueness, continuity wrt $y^0$ in uniform norm and $1-var.$
    \item \textit{local uniqueness}: $V$ locally $Lip^1$ $\Rightarrow$ esistence and uniqueness up to explosion time. If no explosion (e.g. linear growth), global unique solution.
    \item \textit{uniform continuity}: $V^i$ globally $Lip^1$ $\Rightarrow$ uniform continuity wrt $y_0,\ V,\ x$.
    \item \textit{uniform local continuity}: $V^i$ locally $Lip^1$, linear growth $\Rightarrow$ continuity wrt $y_0,\ V,\ x$, with estimates depending on $|x|_{1-var},\ |y_0|,\ V.$
    \item \textit{1-var continuity}: $V^i$ globally $Lip^1$ $\Rightarrow$ $1-var$ continuity wrt $y_0,\ V,\ x$.
    \item \textit{local 1-var continuity}: $V^i$ locally $Lip^1$, linear growth $\Rightarrow$ $1-var$ continuity wrt $y_0,\ V,\ x$, with estimates depending on $|x|_{1-var},\ |y_0|,\ V.$
\end{itemize}

That is, boundedness or linear growth implies global existence, continuity implies existence up to explosion, global $Lip^1$ implies global uniqueness and continuity (uniform and in $1-var$), local $Lip^1$ implies uniqueness up to explosion and continuity (uniform and in $1-var$).
